# Data {#data}

## Source of data

There is a lack of publicly available, representative datasets for the 
evaluation of data linkage processes [@ferrante2012]. "Gold-standard" linked 
datasets are few and, due to confidentiality laws, sharing seldom possible.

This study used datasets derived from data sampled from Wikimedia's publicly 
available Wikidata database [@wikidata]. A sample of records of Wikidata 
entities classified as human was extracted and processed using scripts made 
available by the UK Ministry of Justice's data linkage team [@splink_syn2021]. 
The extracted records comprised data belonging to individuals with a date of 
death between 1953 and 2000 (inclusive).

Research ethics approval for the use of this data is discussed in Appendix 
\@ref(ethics).

## Processed datasets

The supplied linkage evaluation dataset was processed to retain only 
individuals recorded as holding UK citizenship at any point in time 
and having at least one recorded given name and family name. The 
dataset comprised 20,304 records representing distinct individuals. 
The processed dataset consisted of 8 fields, detailed 
in Table \@ref(tab:fieldlist).

```{r fieldlist}

field_list <- data.frame(`field name` = c("unique entity identifier",
                                          "first name",
                                          "family name",
                                          "day of birth",
                                          "month of birth",
                                          "year of birth",
                                          "place of birth",
                                          "postcode",
                                          "gender",
                                          "occupation"),
                         `missing values (%)`= c(0,
                                                 0,
                                                 0.7,
                                                 5.3,
                                                 0,
                                                 0,
                                                 0,
                                                 0,
                                                 1,
                                                 29.6),
                         notes = c("Assigned by Wikidata",
                                   rep("", 5),
                                   "Randomly assigned if not present in Wikidata source",
                                   "Valid UK postcodes assigned based on, sometimes randomly assigned, place of birth",
                                   rep("", 2)))

knitr::kable(field_list, 
             align = "lrl",
             col.names = c("field name", 
                           "missing values (%)", 
                           "notes"),
             digits = 1,
             booktabs = TRUE,
             caption = 'List of fields and percentage of missing values in study dataset.') |> 
  kableExtra::kable_styling(latex_options = "striped", bootstrap_options = "striped")

```

## Test datasets

Many man-made and natural phenomena are distributed according to a 
power-law distribution, this includes frequency of patient attendance 
at emergency departments and other out-of-hours care facilities 
[@burton2018; @burton2022]. We therefore chose to duplicate the the 
study dataset records according to a power-law distribution with 
scaling parameter 3 up to a varying maximum number of duplicates 
per source record.

The attributes of the duplicated records were also corrupted using
mechanisms dependent on the type of attribute and with 
varying probabilities based on those seen in real world datasets [@tromp2011; @hong2013]. 
This corruption aimed to  mimic the real-world data quality issues 
introduced in data capture processes. The mechanisms are described in 
Table \@ref(tab:corruption_types).

```{r corruption_types}

field_list <- data.frame(field = c("first name",
                                   "family name",
                                   "date of birth",
                                   "gender"),
                         mechanism = c("QWERTY typographical error",
                                       "QWERTY typographical error",
                                       "Number pad typographical error",
                                       "Swapped"),
                         base_probability = c(5.0,
                                              5.0,
                                              1.0,
                                              0.5)
)

knitr::kable(field_list, 
             align = "lrl",
             col.names = c("Attribute", 
                           "corruption mechanism", 
                           "Base case probability of corruption (%)"),
             booktabs = TRUE,
             caption = 'Attributes, their corruption mechanism and base case probability of corruption.') |> 
  kableExtra::kable_styling(latex_options = "striped", bootstrap_options = "striped")

```
